Determining sequence-dependent DNA oligonucleotide hybridization and dehybridization mechanisms using coarse-grained molecular simulation, Markov state models, and infrared spectroscopy

### 2021-07-31 response_letter/response_letter.tex

``The effect of the applied bias upon the thermodynamics can be rigorously corrected for using standard reweighting techniques \citep{Gallicchio2005TemperaturePaths, Souaille2001ExtensionCalculations, Shirts2008StatisticallyStates, Kumar1992THEMethod}. \rood{Rigorous elimination of the bias in the kinetics is critical for the construction of robust and reliable Markov state models that reflect the true system dynamics and are uncontaminated by any residual effects of the biasing potentials used to induce good sampling. A number of approaches to correct the kinetics are also available, including Girsanov reweighting, transition-based reweighting analysis (TRAM), dynamic histogram analysis method (DHAM), and their derivatives~\citep{Prinz2011OptimalDynamics, Chodera2011DynamicalTemperatures, Stelzl2017DynamicSimulations, Donati2017GirsanovModels, Donati2018GirsanovSimulations, Quer2018ANCOORDINATES} \citep{Mey2014XTRAM:States}. The application of these methods under the conditions of high bias necessary for good sampling can, however, present challenges for numerical convergence.} A number of coarse-grained DNA force fields have been developed that enable direct observation of these events over microsecond time scales via unbiased coarse-grained molecular dynamics simulations, \citep{Romano2013DNADependence, Hinckley2013AnHybridization, Maciejczyk2014DNAModel, Markegard2015, Dans2016MultiscaleDNA} which, up to an acceleration factor associated with the smoothing of the underlying free energy landscape inherent to the coarse-graining procedure, can preserve a faithful model of the unbiased dynamics and associated pathways. These models have previously been used to study biological phenomena such as nucleosome dynamics\citep{Lequieu2016Tension-dependentUnwrapping, Lequieu2017InSliding} and transcription factor binding \cite{Terakawa2015P53Simulations, Tan2018DynamicData} as well as nanotechnology applications such as strand displacement \citep{Srinivas2013OnDisplacement, Haley2020DesignDisplacement} and DNA origami \citep{Snodin2019Coarse-grainedOrigami, Doye2020TheOrigami}. \rood{In this study, we choose to employ a coarse-grained model for DNA \citep{Hinckley2013AnHybridization} that is sufficiently inexpensive to enable the collection of sufficient volumes of unbiased simulation trajectories and adequately sample configurational space that we do not need to appeal to biasing strategies to enhance convergence nor apply any \textit{post hoc} corrections to the thermodynamics or kinetics.}''

@@

``A single pair of self-complementary sequences were placed in a cubic periodic box with side length 7.8 nm corresponding to a single-strand concentration of \rood{7 mM}. \rood{This concentration is only 3.5$\times$ larger than the 2 mM concentration employed in our experimental analyses (Section 2.2).}''

(III) Finally, we deliberately chose to explore the melting temperature as the state point at which the hybridized and dehybridized populations are in balance and sought to probe the mechanism and kinetics of the hybridization/dehybridization transitions under these equilibrium conditions corresponding to no net thermodynamic preference for either state (i.e., minimally driven conditions). A favorable side effect of this choice was indeed that we observe approximately equal numbers of hybridization and dehybridization events in a single (sufficiently long) simulation, whereas temperatures away from $T_\mathrm{melt}$ would favor one over the other. It is an interesting question to consider how the mechanism and kinetics of hybridization/dehybridization may change with temperature. Exploring small excursions either side of $T_\mathrm{melt}$ may still be possible using unbiased simulations, but we concur with the reviewer that biased sampling would be required to observe sufficient numbers of hybridization (dehybridization) events at large excursions above (below) $T_\mathrm{melt}$.

In response to this point, we have modified text in the terminal paragraph of the Conclusions on p.~38 observing that this may be an interesting point for follow-up study:

``Going forward, we will extend this work to discern more general trends in sequence-dependent hybridization/dehybridization for a wider range of oligomer sequences and motivate strategies for experimental comparisons. \rood{We also suggest that the MSM approach followed in this work, possibly coupled with biased sampling and reweighting techniques \citep{Wu2014StatisticallyStates, Scherer2015PyEMMAModels}, may be well-suited to expose changes in the hybridization/dehybridization mechanism and kinetics as a function of temperature.}''

@@

\textcolor{blue}{Very small points:
+ p16 line 3: "Although the 3SPN.2 model reproduces melting temperatures relatively well, we observed a systematic 4 K under-prediction relative to experiment and so we apply a universal (+4) K corrective calibration to our computational results". This sentence confused me at first. I presume you are just shifting your simulated kd slow curves to better match experiment, rather than shifting the data so that the melting temperatures agree. You may want to be more explicit as I read it as the latter at first. (Note one shouldn't expect the melting temperatures to agree as very different concentrations are being used in simulation and experiment, and one also has to correct for finite-size effects when estimating melting points from single duplex simulations (https://doi.org/10.1088/0953-8984/22/10/104102)}

\textbf{Author Reply:}   (I) We regret our lack of clarity in our choice of wording. Yes, we simply shifted the $k_d$ curves by a universal temperature shift, determined empirically to be +4 K, that produced the optimal match with the experimental data.

We have also reworded the unclear wording regarding the temperature correction in Section 3.1 p.~17 to now read:

@@

``Although the 3SPN.2 model reproduces melting temperatures relatively well, we observed a systematic 4 K under-prediction relative to experiment and so we apply a \rood{universal (+4) K corrective temperature shift} to our computational results.''

(II) We regret our erroneous reporting of a 7 M rather than 7 mM concentration as detailed in our previous response. As we have now clarified, our simulations and experiments are conducted at similar concentrations (7mM and 2 mM, respectively). We agree that we still cannot expect perfect agreement in the melting temperatures due to the residual small  difference in concentrations, but likely this is small in comparison to discrepancies introduced by the approximations inherent in the coarse-grained molecular model.

(III) We greatly appreciate this comment and suggestion to apply a fluctuation correction to our data in order to correct for artifacts associated with finite-size effects following the approach detailed in Ref.~\citep{Ouldridge2010ExtractingSimulations}. We have reanalyzed our slow repsonse data under the homodimer correction and this has pleasingly resulted in a slightly improved agreement between our simulations and experiment. The revised figure is reproduced below.

\begin{figure}[ht!]
	\begin{center}
        \includegraphics[width=0.8\textwidth]{../Fig1.pdf}
        \caption*{\textbf{Figure 1.} Experimental measurements and computational predictions of slow and fast at T-jump IR responses. Results are reported in terms of the final T-jump temperature. (a) The experimental and simulated slow rate constants $k_d^\mathrm{slow}$ corresponding to duplex dissociation over long time scales. \rood{We apply a fluctuation correction proposed by Ouldridge et al.\citep{Ouldridge2010ExtractingSimulations} to account for finite-size effects and ensure an asymptotically correct dimer fraction at long times.} (b) The experimental and simulated fast rate constant $k_d^\mathrm{fast}$ corresponding to terminal base-pair fraying on short time scales. The simulation results are corrected by a sequence-independent scaling factor that corrects for a 10$\times$ acceleration of the slow dissociation dynamics and 120$\times$ acceleration of the fast fraying dynamics. The simulated temperature in all cases is subjected to a (+4) K corrective calibration to account for an observed systematic under-prediction of the melting temperature by the 3SPN.2 model}
        \label{fig:relaxation-comparison}
	\end{center}
\end{figure}

@@

In light of this modification, we have also added the following text to the Section 3.1 on p.~16:

\rood{To correct for finite-size effects associated with the simulation of a single duplex within a periodic simulation box and ensure that we will observe asymptotically correct behavior at long times, we apply the fluctuation correction proposed by Ouldridge et al.\citep{Ouldridge2010ExtractingSimulations} to our predictions for $f_\mathrm{hybridized}$ bulk yields. We correct the values observed in our simulations $f_\mathrm{hybridized}^\mathrm{sim}$ to the bulk corrected result $f_\mathrm{hybridized}$ as,}

@@

``We first compare the thermodynamic predictions for the equilibrium macrostate probabilities \rood{made by the MSM models fitted at the sequence melting temperatures} to those of the nearest neighbor (NN) model as a popular empirical model of DNA hybridization thermodynamics.''

In the first sentence of the caption of Fig.~2, we have added:

``Thermodynamic and kinetic predictions of the sequence-specific MSMs \rood{fitted at the sequence melting temperatures} for AT-all, GC-end, GC-core, and GC-mix.''

In the first sentence of the caption of Fig.~3, we have added:

@@

\textbf{Author Reply:}    (I) We share the reviewer's concerns with using ``slithering'' to describe a process that is more appropriately referred to as ``defect diffusion'', ``inchworm mechanism'', or ``out-of-register diffusion''.

In Section 2.1.2 on p.~9, we have modified the text to replace ``slithering'' with ``out-of-register diffusion'':

``An energy disconnectivity graph-based approach was used to interrogate the differences in hybridization rates and mechanisms between GGGGGG and GCGCGC hexamers to reveal strong deviations from ``all-or-nothing'' behaviors and the importance of zippering and \rood{out-of-register diffusion mechanisms.~\citep{Xiao2019}''}

The only other place we use the term is in Section 3.4 on p.~29 where we refer to work conducted in Ref.~\citep{Xiao2019}. Here we choose to maintain the term ``slithering'' to maintain consistency with the mechanism nomenclature adopted in this paper, but place it in quotation marks to indicate that we are quoting from that paper and it is not a term that we would necessarily favor.

``Out-of-register states for 5$^\prime$-GCGCGC-3$^\prime$ hexamers were identified as deep kinetic traps along the hybridization pathway and ``slithering'' through these states did not provide a significant hybridization pathway compared to an alternative ``zippering'' mechanism. (In contrast, out-of-register slithering and in-register zippering served as two parallel pathways for hybridization of 5$^\prime$-GGGGGG-3$^\prime$.)''

@@

(II) We also appreciate the reviewer pointing out the incorrect author list in ref 52, and we have fixed the issue.

\textcolor{blue}{4. P15: I am a bit surprised that a separation of 1.3nm between the central base pairs is enough to guarantee strand separation, given that frayed end can separate by 1.3nm when the core is intact. Can the authors verify that all their duplexes had truly separated at this point?}

\textbf{Author Reply:}   We motivated our choice of this distance as that beyond which the strands are effectively non-interacting, but also verified that our estimates of $k_d^{slow}$ are robust to choices of this cutoff in the range 0.8--2.5 nm. As an illustrative example, we demonstrate in \textbf{Figure R1} below that our extracted values of $k_d^{slow}$ are robust to changing the cutoff from 1.3 nm to 2.0 nm.

\begin{figure}[ht!]
	\begin{center}
        \includegraphics[width=\textwidth]{../FigR1.pdf}
        \caption*{\textbf{Figure R1.} Comparing the experimental slow response when increasing the cutoff parameter from 1.3 nm to 2.0 nm}
        \label{fig:compare_cutoffs}
	\end{center}
\end{figure}

@@

We have modified the relevant text in Section 3.1 on p.~15 to read:

``First, we tracked the slow response corresponding to duplex dissociation in our simulations by compiling the distribution of times at which both of the central base pairs first separate to a distance of 1.3 nm starting from an initial fully hybridized duplex. This cutoff was selected as the distance beyond \rood{which} the strands are effectively non-interacting and define the dissociated state. \rood{We verified that our results were robust to choices of this cutoff over the range 0.8-2.5 nm.}''

@@

\textcolor{blue}{5. The authors refer to "exponential" increase in Fig. 1. However, it's hard to see that an increase is exponential by eye unless the data is plotted on a log scale.}

\textbf{Author Reply:}   We agree that a log-linear plot would better expose these trends. We have replotted Fig.~1a and Fig.~S4 on log-linear axes and reproduce these below.

\begin{figure}[ht!]
	\begin{center}
        \includegraphics[width=1.0\textwidth]{../Fig1.pdf}
        \caption*{\textbf{Figure 1.} Experimental measurements and computational predictions of slow and fast at T-jump IR responses. Results are reported in terms of the final T-jump temperature. (a) The experimental and simulated slow rate constants $k_d^\mathrm{slow}$ corresponding to duplex dissociation over long time scales. \rood{We apply a fluctuation correction proposed by Ouldridge et al.\citep{Ouldridge2010ExtractingSimulations} to account for finite-size effects and ensure an asymptotically correct dimer fraction at long times.} (b) The experimental and simulated fast rate constant $k_d^\mathrm{fast}$ corresponding to terminal base-pair fraying on short time scales. The simulation results are corrected by a sequence-independent scaling factor that corrects for a 10$\times$ acceleration of the slow dissociation dynamics and 120$\times$ acceleration of the fast fraying dynamics. The simulated temperature in all cases is subjected to a (+4) K corrective calibration to account for an observed systematic under-prediction of the melting temperature by the 3SPN.2 model}
        \label{fig:relaxation-comparison_2}
	\end{center}
\end{figure}

@@

\clearpage
\newpage

% Fig. S4
\begin{figure}[ht!]
	\centering
    \includegraphics[width=.8\textwidth]{../FigS4.pdf}
   \caption*{\textbf{Figure S4.} Exponential fits for both slow (dissociation) and fast (fraying) response for all four sequences during ``computational T-jump'' experiments. From 120 independent 1 $\mu$s simulations, we compiled the slow response data by recording the fraction of sequences with both central Watson-Crick base pairs intact as a function of time, and the fast response data as the fraction of sequences with both terminal Watson-Crick base pairs intact as a function of time. We define a Watson-Crick base pair to be intact if the centers-of-mass of the two comlementary bases lie within a linear distance of 1.3 nm. We extracted our computational estimate of $k_d^\mathrm{fast}$ by fitting a decaying exponential to the fraction of bound A:T termini as a function of time $f_\mathrm{unfrayed}(t) = \exp(-k_d^\mathrm{fast}t)$. Similarly, we extracted our computational estimate of $k_d^\mathrm{slow}$ by fitting a decaying exponential to the fraction of hybridized sequences as a function of time $f_\mathrm{hybridized}(t) = \exp(-k_d^\mathrm{slow}t)$. \rood{We report within the figure legend to each panel the coefficient of determination $R^2$ for a least squares linear fit of the model to the data in log space (i.e., $\log \left( f \right) = -k_d t$) and data are plotted on log-linear axes to facilitate visual comparison of the fits. In all cases we observe excellent fits of the models to the data with all $R^2$ > 0.88 except for the slow response at the the lowest temperature T$_m$-5K where dissociation events are sparse.}}
    \label{fig:SIFig4_2}
\end{figure}

@@

\clearpage
\newpage

\textcolor{blue}{6. Do the authors have any data on the scaling of k\_hybridization with T? In experiment, rates appear to increase with temperature (albeit with a much weaker dependence than de-hybridization rates). Simple "nucleation and zipper" models would predict the opposite, as does the oxDNA model (see ref 52).}

\textbf{Author Reply:}   This is an interesting question and are, in fact, considering embarking upon some additional work in this vein. Rigorous engagement of this question would require us to fit MSMs at a variety of temperatures in order to compare the pathways, mechanisms, and rates for hybridization and dehybridization. In the present work, we have only constructed MSMs at the melting temperature of each sequence and the collection of sufficient data to parameterize MSMs at other temperatures would require both (i) a substantial investment of time to collect the $\sim$1 ms of data required to fit robust MSMs and (ii) the use of biasing techniques in order to adequately sample hybridization (dehybridization) events at large temperature excursions above (below) the melting temperature. As such, we do not have the necessary data to engage this question at the present time.

Reviewer \#2 raised a related point in their Comment \#2 and in response to both of these points, we have made the following modification to the terminal paragraph of the Conclusions on p.~38 observing that this may be an interesting follow-up study:

``Going forward, we will extend this work to discern more general trends in sequence-dependent hybridization/dehybridization for a wider range of oligomer sequences and motivate strategies for experimental comparisons. \rood{We also suggest that the MSM approach followed in this work, possibly coupled with biased sampling and reweighting techniques \citep{Wu2014StatisticallyStates, Scherer2015PyEMMAModels}, may be well-suited to expose changes in the hybridization/dehybridization mechanism and kinetics as a function of temperature.}''

@@

\textcolor{blue}{7. The authors could explain a bit more carefully (perhaps in the SI) how they handle the multiple nearest-neighbour states that correspond to their microstates (eg. F4 clearly involves more than one NN configuration).}

We have modified the text in the ``Nearest neighbor model of duplex thermodynamics'' text in the Supporting Information to better describe for the reader how we performed this calculation. Specifically, we have appended the following text at the end of this discussion:

``\rood{In applying the NN model to each macrostate, we made the simplifying assumption that the ensemble of microstates constituting each macrostate could be represented by a single pattern of Watson-Crick base pairing that are schematically illustrated in Fig.~2a. In reality, each macrostate is composed of an ensemble of microstates that may have slightly different base pairing patterns and explicitly averaging over all of these microstates may change the predictions of the NN model. Recalculating the NN predictions by explicitly averaging over all microstates in the macrostate leads to changes in the $\Delta F = F - F_H$ values reported in Fig.~3 (i.e., the free energy of each macrostate relative to the hybridized state) of $<$1 kJ/mol. As such, representing each macrostate by a single representative microstate and base pairing pattern is a highly accurate simplifying approximation for the purposes of this calculation.}''

@@

\textcolor{blue}{8. I don't understand how the population of F4 could be >50\% that of H for the GC-core system, when the free energy of F4 is reported as > 5kJ/mol. My back-of-the-envelope maths predicts a ratio of 8:1 for a 5kJ/mol gap. Incidentally, 8:1 is roughly the ratio I got from running the configurations through NUPACK.}

\textbf{Author Reply:}  We apologize for the confusion and have resolved the origin of this apparent inconsistency to two factors: (i) a unit conversion error on our part and (ii) our choice to present the bar plots in Fig.~2b on a log axis.

(I) First, we made an (embarrassing) unit conversion error in reporting the $\Delta F$ value for the F4 state of the GC-core sequence from our MSM model. Specifically, we did not convert the $\Delta$F from units of $kT$ at $T_m$ = 324 K to kJ/mol, resulting us to erroneously report a value of 5.1 kJ/mol instead of 5.1 $kT$ = 13.7 kJ/mol. Using this correct $\Delta$F value we see agreement between the reported free energy in Fig.~3 and populations reported in Fig.~2. We can demonstrate this by performing the same back-of-the-envelope verification proposed by the reviewer:

### 2021-07-31 response_letter/response_letter.tex

\begin{itemize}
\item $G_H$ = $G_D$ = 0 (adopting this as the arbitrary zero of free energy and asserting that the sequence is at its melting temperature so the hybridized H and dissociated D states are equiprobable)
\item $G_{F4}$ = 5.1 $kT$ at $T_m$ = 324 K = 13.7 kJ/mol
\item $f_H$ + $f_D$ + $f_F$ = 1 (the fractional populations sum to unity)
\item $f_X$ = $\exp (-\beta G_X$) / $Z$ ($Z$ is the partition function that imposes normalization)
\end{itemize}

We can solve this system of equations to yield $f_{F4}$ = 0.0030 that is in excellent agreement with the MSM predictions of $f_{F4}$ = 0.0028 that we report in Fig.~2b.

We have updated Figure 3 with the correct $\Delta$F as shown below.

\begin{figure}[ht!]
	\begin{center} 
        \includegraphics[width=140mm, scale=1]{../Fig3.pdf}
        \caption*{Figure 3. Comparison of the macrostate free energy predictions of the MSMs and nearest neighbor (NN) thermodynamic model \rood{at the sequence melting temperatures}.~\citep{SantaLucia1998AThermodynamics, Santalucia2004TheMotifs} (a) Free energies of each macrostate relative to the hybridized state $\Delta F = F - F_H$. We define the hybridized state H to possess a free energy of zero and take care to only compare relative free energies (i.e., $\Delta F$) between the MSM and NN model. (b) Discrepancy between the macrostate relative free energy predictions $\Delta \Delta F = \Delta F^\mathrm{MSM} - \Delta F^\mathrm{NN}$ of the MSM relative to the NN model. The MSM tends to predict higher relative free energies (i.e., lower occupancy probabilities) relative to the hybridized state H compared to the NN model.}
        \label{fig:NN_table}
	\end{center}
\end{figure}

We have also updated the second paragraph of Section 3.3 on p.~23 as follows:

``As illustrated in Fig.~3, we see that the MSM tends to predict higher free energies for \rood{all macrostates relative to the H state compared to the NN model}. Although the trends are in qualitative agreement, what is the root of the quantitative discrepancy of the MSM and NN models in the predicted relative stabilities?''

We recognize that this was a significant miscalculation, and we have re-checked all free energy results to ensure that (i) we did not make a similar error for any of the other reported values in Fig.~3 or anywhere else and (ii) that this error did not propagate through to any subsequent calculations, results, or conclusions within the remainder of the manuscript.

(II) We also note that our choice of a logarithmic y-axis for the bar plots in Fig.~2b that make the bar associated with the F4 state appear nearly 50\% of the height of the H and D bars for GC-core. We made this plotting choice in order to make visible on these bar plots the small state populations associated with the shifted and frayed states and noted the percentage occupancies on the bars themselves. To help eliminate this confusion, we have modified the portion of the Fig.~2 caption referring to this panel to read:

@@

``(b) \textbf{Thermodynamics.} Histograms reporting the number of the $10^7$ total frames within the 1 ms of simulation trajectories observed to occupy each of the seven macrostates, corresponding to our numerical estimates of the equilibrium occupancy probabilities. Uncertainties are calculated across 100 MSMs using a Bayesian MSM estimation are reported for each bar and are very small compared to the total counts. \rood{Values are reported on a log y-axis to make the small populations of the shifted and frayed states visible.}''

\textcolor{blue}{9. Is the MSM model explicitly constrained to obey detailed balance? The numbers that come out seem consistent with detailed balance - forward and backward trajectories at the same T essentially look the same (they're equally likely to go through intermediates). This should be a feature of the MSM, but it's unclear to me whether it's hard-coded or just arises naturally in the coarse-graining.}

\textbf{Author Reply:}    The reviewer raises a good point regarding the assumptions made during MSM construction. We do indeed use a reversible MSM framework that enforces detailed balance. This is an important point which we have now made explicit by modifying the text in Section 2.1.2 on p.~11 to read:

@@

``The 200 microstates comprising each system were coarsened into our terminal macrostate MSM. \rood{Since we are interested in the construction of models of the equilibrium kinetics, we explicitly enforce detailed balance in the construction of the MSM. \citep{Wehmeyer2019IntroductionSoftware}}''

and the text in Section 3.4 on p.~26 to read:

``In addition to thermodynamic stabilities, the macrostate MSM also furnishes quantitative and interpretable predictions of hybridization and dehybridization pathways and mechanisms \rood{at the sequence melting temperatures}. \rood{It should be noted that because we are using a reversible MSM framework, detailed balance is enforced by construction.}''

@@

\textcolor{blue}{Relatedly, some of the discussion of the hybridization and de-hybridization pathways in non-repetitive sequences seems to ignore this fact. In these equilibrium simulations, hybridization trajectories played in reverse should be statistically indistinguishable from de-hybridization trajectories. I don't think anything that's said is wrong, per se, but it might be worth considering how the contrast between "nucleation and zippering" and "fraying-peeling" is consistent with the symmetry in the pathways.}

\textbf{Author Reply:}  This is a nice observation, and we would do well to note this symmetry for the reader. We have added a sentence to this effect in Section 3.6 on p.~32:

@@

\textbf{Author Reply:}   We thank the reviewer for this important observation and illuminating our need to better clarify this potentially ambiguous concept. We have modified the text in Section 3.6 on p.~29 to read:

``The GC-mix sequence is the only one of the four sequences studied that exhibits simple two-state ``all-or-nothing'' behavior.~\citep{Xiao2019, Araque2016LatticeCooperativity, Sikora2013ModelingIntermediates, Sanstead2016} \rood{This characterization is intended to reflect that the mechanistic observation that} GC-mix MSM comprises just two states, the hybridized H and dehybridized D (Fig.~2c), indicating that association and dissociation of the strands proceeds directly without passing through any metastable intermediate states resolvable under the $\tau$=1.2 ns lag time of our MSM. \rood{Indeed, the sequence's propensity to fray in simulation and the role of termini fraying in dissociation experiments \citep{Sanstead2016} indicate that there remain deviations from ``all-or-nothing'' behavior which are not fully captured within the MSM lag time. (We note that ``all-or-nothing'' can also be used, somewhat ambiguously, to refer to a two-state thermodynamic model where the hybridized and dehybridized are separated by a large free energy barrier but that the transitions between them may proceed through a network of metastable states.)}''

We have also removed ``all-or-nothing'' from the experimental methods section as the fitting procedure is more appropriately described as simply ``two-state'' so that the relevant text in Section 2.2.3 on p.~14 now reads:

@@

``FTIR temperature series were measured for each sequence as reported previously,~\citep{Sanstead2016} and the second SVD component along temperature was fit to a \rood{two-state model} to determine the fraction of intact duplex $\theta$ as a function of temperature.''

\textcolor{blue}{11. Related to the above, the references on line 3 of p4 should probably also include Ref 52, where a similar analysis of metastable states in a repetitive sequence is performed (for a different coarse-grained model) and [Flamm, W., Fontana,J.M., Hofacker, I.L. and Schuster,P. (2000) RNA folding at elementary step resolution. RNA, 6, 325-338] where a defect diffusion mechanism is proposed.
}

\textbf{Author Reply:}   We have added these references to those at the top of p.~4.

@@

\textcolor{blue}{12. The reasoning behind why the broader relaxation spectrum is indicative of misaligned states was not that clear to me. Why would "elevated" fraying cause a broader response? Wouldn't it also, then, be broader for CG-core, which has elevated fraying? It seems to me that the argument is more along the lines of the heterogeneity of the configurations leading to heterogeneous dynamics, and hence a broader relaxation profile. But I am no expert on this technique.}

\textbf{Author Reply:}   The reviewer makes a good comment on our choice of language used to described the stretched AT-all response. It is indeed our contention that the heterogeneity of the configurations in the initial ensemble and the fraying dynamics exhibited in each of these states that is the origin of the broadening of the relaxation profile, but our wording here was insufficiently clear. We believe that the reviewer's phrasing ``heterogeneity of configurations leads to heterogeneous dynamics'' expresses this idea much more clearly and we have integrated it into our edits of Section 3.7 on p.~33:

@@

``Nevertheless, the presence of these out-of-register shifted states in the low-temperature equilibrium ensemble prior to the T-jump step may be observable via their influence on the fast T-jump IR response attributable to fraying. Specifically, we hypothesize that the dangling ends and inert tails present in the out-of-register shifted states should promote an \rood{elevated fraying response over the course of the relaxation that is distinct from that of in-register fraying. This heterogeneity of configurations should lead to heterogeneous dynamics, manifested in the observation of a more stretched relaxation over experimental time scales of 70-100 ns.} Analysis of the MSM equilibrium distributions reveals 10.0\% of the equilibrium ensemble to reside in out-of-register shifted states 3S2, 5S2, 3S4, and 5S4 for AT-all, compared to just 0.23\% for GC-end, and 0\% for GC-core and GC-mix. It is our conjecture that a substantial population of out-of-register shifted states in the pre-T-jump AT-all ensemble should be distinguishable from the GC-end, GC-core, and GC-mix as an elongation of the fast relaxation response associated with terminal base fraying.''

\textcolor{blue}{13. How much of the code is being made available? Ideally, the initialization files for the simulations and the code used to perform the MSM analysis would be publicly available. Additionally, it would be good to publish the raw experimental data and the processing tools used to obtain the final results.}

\textbf{Author Reply:}   We have uploaded input files and analysis scripts to Github (\url{https://github.com/mrjoness/Oligo-MSMs}) and 45 GB of featurized trajectories and experimental data to Zenodo (\url{https://doi.org/10.5281/zenodo.5140040}). The links to both of these resources are now noted in the Supporting Information statement on p.~38:
